---
title: "Practical Machine Language Project"
author: "Vasudha Upadhyaya"
date: "Wednesday, February 11, 2015"
output: html_document
---


#### Synposis

This report analyzes to quantify how well an individual will perform  a particular activity. In this project, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. People were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). For the report to be  reproducible, I have used set.seed(). For predictions I have used random forest becuase the results are mostly accurate.

### Basic settings

```{r}
echo = TRUE  # Always make code visible
options(scipen = 1)  # Turn off scientific notations for numbers

library(caret)
library(kernlab)
library(randomForest)
setInternet2(TRUE)
```

## Data Cleaning and Processing
The most latest data -traing and testing  are downloaded and saved as pml-traing.csv and pml-testing.csvWe read the generated csv file.To clean the data, the first row index and all colomuns with NA were removed. . 

```{r}
rm(list = ls()) 
# Read cleaned training and testing data 

train_url<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(train_url,destfile="pml-training.csv")
download.file(test_url,destfile="pml-testing.csv")

train<-read.csv("pml-training.csv",na.strings=c("NA",""))
test<-read.csv("pml-testing.csv",na.strings=c("NA",""))

training<-train[,-c(1:7)]
set.seed(333)
```

###Building data sets for training 

Using 70% for training and 30% for Cross Validation. None generated for testing since that set is already provided.

```{r}
inTrain<-createDataPartition(training$classe, p=0.75, list=FALSE)
training_train<-training[inTrain,]
training_test<-training[-inTrain,]

dim(training_train)
dim(test)

training_train<-as.data.frame(training_train)
clean_train<-training_train[,colSums(is.na(training_train))==0]
clean_test<-training_test[,colSums(is.na(training_test))==0]


```
#### Define cross-validation experiment

```{r}


fitControl = trainControl( method = "cv", number = 4)
# Perform the cross validation
validation <- train(classe ~ ., data = clean_train, method = "rf", 
  trControl = fitControl)
bestTune<-validation$bestTune$mtry

```

### Data Analysis
```{r}
validation
```

###Build random forest model with full training model

```{r}
rForest = randomForest(classe ~ ., data = clean_train,
                            mtry = bestTune)
predictTraining = predict(rForest)
table(predictTraining, clean_train$classe)
```

Predict testing data
```{r}
predictTesting = predict(rForest, newdata = test)
predictTesting
```

Write the Prediction to files
```{r}
# Function to write a vector to files
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_", i ,".txt")
    write.table(x[i], file = filename, quote = FALSE,
                row.names = FALSE, col.names = FALSE)
  }
}
# Call the function
pml_write_files(predictTesting)
```